---
title: "rcca"
author: ""
date: "2021/10/26"
output: 
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Library
```{r}
library(mixOmics)
library(dplyr)
```

## Data
```{r}
load("../../input/3omics_combined/SixHosp_5groups_319subjects_3omics.RData")
immune <- data.frame(immune)
ck <- data.frame(ck)
immune <- immune[!is.na(immune$T),]
ck <- ck[match(rownames(immune), rownames(ck)),]
meta_info <- meta_info[match(rownames(immune), meta_info$sample_FACS),]
```

## rCCA: NC vs SCD1
```{r}
groups <- c("NC", "SCD1")
meta_sel <- meta_info[meta_info$Group %in% groups,]
meta_df <- meta_sel

meta_df <- meta_df[order(meta_df$Group),]
int_id <- meta_df$sample_FACS
ck_df <- ck[match(int_id, rownames(ck)),]
immune_df <- immune[match(int_id, rownames(immune)),]

X <- log2(ck_df + 0.0001)
Y <- log2(immune_df + 0.0001)
```

### Estimation of penalisation parameters (shrinkage method)
```{r}
rcc.shrink <- rcc(X, Y, ncomp = 2, method = 'shrinkage')
plot(rcc.shrink)
```

#### Sample Plots
```{r, fig.width=7, fig.height=5}
plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, rep.space = "XY-variate",
          legend = TRUE, title = 'rCCA XY-space')

plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, 
          legend = TRUE, title = 'each subspace')
```

#### Variable Plots
```{r, fig.width=11, fig.height=9}
#cutoff = 0.5 so that only the variables outside the small circle of radius 0.5 are represented
plotVar(rcc.shrink, comp = 1:2, cutoff = 0.5, var.names = c(TRUE, TRUE),
        cex = c(4, 4), title = 'rCCA comp 1 - 2, corr 0.5')

cim(rcc.shrink, comp = 1:2, xlab = "cytokine", ylab = "immune", 
    margins = c(5, 6))
```



## rCCA: SCD1 vs aMCI
```{r}
groups <- c("SCD1", "aMCI")
meta_sel <- meta_info[meta_info$Group %in% groups,]
meta_df <- meta_sel

meta_df <- meta_df[order(meta_df$Group),]
int_id <- meta_df$sample_FACS
ck_df <- ck[match(int_id, rownames(ck)),]
immune_df <- immune[match(int_id, rownames(immune)),]

X <- log2(ck_df + 0.0001)
Y <- log2(immune_df + 0.0001)
```

### Estimation of penalisation parameters (shrinkage method)
```{r}
rcc.shrink <- rcc(X, Y, ncomp = 2, method = 'shrinkage')
plot(rcc.shrink)
```

#### Sample Plots
```{r, fig.width=7, fig.height=5}
plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, rep.space = "XY-variate",
          legend = TRUE, title = 'rCCA XY-space')

plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, 
          legend = TRUE, title = 'each subspace')
```

#### Variable Plots
```{r, fig.width=11, fig.height=9}
#cutoff = 0.5 so that only the variables outside the small circle of radius 0.5 are represented
plotVar(rcc.shrink, comp = 1:2, cutoff = 0.5, var.names = c(TRUE, TRUE),
        cex = c(4, 4), title = 'rCCA comp 1 - 2, corr 0.5')

cim(rcc.shrink, comp = 1:2, xlab = "cytokine", ylab = "immune", 
    margins = c(5, 6))
```


## rCCA: aMCI vs AD1
```{r}
groups <- c("aMCI", "AD_1")
meta_sel <- meta_info[meta_info$Group %in% groups,]
meta_df <- meta_sel

meta_df <- meta_df[order(meta_df$Group),]
int_id <- meta_df$sample_FACS
ck_df <- ck[match(int_id, rownames(ck)),]
immune_df <- immune[match(int_id, rownames(immune)),]

X <- log2(ck_df + 0.0001)
Y <- log2(immune_df + 0.0001)
```

### Estimation of penalisation parameters (shrinkage method)
```{r}
rcc.shrink <- rcc(X, Y, ncomp = 2, method = 'shrinkage')
plot(rcc.shrink)
```

#### Sample Plots
```{r, fig.width=7, fig.height=5}
plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, rep.space = "XY-variate",
          legend = TRUE, title = 'rCCA XY-space')

plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, 
          legend = TRUE, title = 'each subspace')
```

#### Variable Plots
```{r, fig.width=11, fig.height=9}
#cutoff = 0.5 so that only the variables outside the small circle of radius 0.5 are represented
plotVar(rcc.shrink, comp = 1:2, cutoff = 0.4, var.names = c(TRUE, TRUE),
        cex = c(4, 4), title = 'rCCA comp 1 - 2, corr 0.4')

cim(rcc.shrink, comp = 1:2, xlab = "cytokine", ylab = "immune", 
    margins = c(5, 6))
```


## rCCA: AD1 vs AD2
```{r}
groups <- c("AD_1", "AD_2")
meta_sel <- meta_info[meta_info$Group %in% groups,]
meta_df <- meta_sel

meta_df <- meta_df[order(meta_df$Group),]
int_id <- meta_df$sample_FACS
ck_df <- ck[match(int_id, rownames(ck)),]
immune_df <- immune[match(int_id, rownames(immune)),]

X <- log2(ck_df + 0.0001)
Y <- log2(immune_df + 0.0001)
```

### Estimation of penalisation parameters (shrinkage method)
```{r}
rcc.shrink <- rcc(X, Y, ncomp = 2, method = 'shrinkage')
plot(rcc.shrink)
```

#### Sample Plots
```{r, fig.width=7, fig.height=5}
plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, rep.space = "XY-variate",
          legend = TRUE, title = 'rCCA XY-space')

plotIndiv(rcc.shrink, comp = 1:2, ind.names = meta_df$Group, 
          group = meta_df$Group, 
          legend = TRUE, title = 'each subspace')
```

#### Variable Plots
```{r, fig.width=11, fig.height=9}
#cutoff = 0.5 so that only the variables outside the small circle of radius 0.5 are represented
plotVar(rcc.shrink, comp = 1:2, cutoff = 0.5, var.names = c(TRUE, TRUE),
        cex = c(4, 4), title = 'rCCA comp 1 - 2, corr 0.5')

cim(rcc.shrink, comp = 1:2, xlab = "cytokine", ylab = "immune", 
    margins = c(5, 6))
```
