---
title: "02_Normalization_DimensionalityReduction_Clustering"
author: ""
date: "2022/10/18"
output: 
  html_document: 
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

We then apply the standard workflow to each sample separately:

normalisation
variance modelling
dimensionality reduction
clustering

## Load library
```{r}
library(SingleCellExperiment)
library(scuttle)
library(scater)
library(BiocParallel)
library(scran)
library(batchelor)
library(bluster)
library(cowplot)
```

## Normalization

The normalized values are used for visulization, clustering and dimensionality reduction.

```{r}
# Parameters
this.path <- "../Results/02_normalization_and_dimensionality_reduction_clustering"
bp.params <- MulticoreParam(workers = 20)
set.seed(1234)
dir.create(this.path)

# Import sce object after quality control
sce <- readRDS("../Results/01_preprocess_and_quality_control/SCE_filtered_cellQC_3MAD_featureQC_above20cells.RDS")

# Split sce by column data into list
sample.sce <- levels(factor(colData(sce)$Sample))
all.sce <- lapply(sample.sce, function(x) sce[, colData(sce)$Sample == x])
names(all.sce) <- sample.sce

#--- normalization ---#
# logcounts: log-transformed normalized expression values
all.sce <- lapply(all.sce, logNormCounts)

#--- variance-modelling ---#
# model variance with modelGeneVar()
# find top 2000 highly variable genes (HVGs) with getTopHVGs()
all.dec <- lapply(all.sce, modelGeneVar)
all.hvgs <- lapply(all.dec, getTopHVGs, n=2000)
```

## Dimensionality reduction
```{r}
#--- dimensionality-reduction ---#
# use runPCA()
# then compute embeddings with runTSNE() and runUMAP()
all.sce <- mapply(FUN=runPCA,
          x=all.sce,
          subset_row=all.hvgs,
          MoreArgs=list(ncomponents=30,
                BSPARAM=RandomParam()),
          SIMPLIFY=FALSE)

all.sce <- lapply(all.sce, runTSNE, dimred="PCA")
all.sce <- lapply(all.sce, runUMAP, dimred="PCA")

#--- clustering ---#
# cluster each sample separately
for (n in names(all.sce)) {
    g <- buildSNNGraph(all.sce[[n]], k=10, use.dimred='PCA')
    clust <- igraph::cluster_walktrap(g)$membership
    all.sce[[n]]$label  <- factor(clust)
}
# saveRDS(all.sce, file = paste0(this.path, "/SCE_list_PCA_TSNE_UMAP_clustering.RDS"))
# all.sce <- readRDS(paste0(this.path, "/SCE_list_PCA_TSNE_UMAP_clustering.RDS"))
```

### Data prepared for batch effect correlation
To prepare for the batch correction:
We subset all all the samples to the common “universe” of features (Ensembl gene annotation). 

Then we rescale each batch to adjust for differences in sequencing depth between batches. The multiBatchNorm() function recomputes log-normalized expression values after adjusting the size factors for systematic differences in coverage between SingleCellExperiment (SCE) objects. (Size factors only remove biases between cells within a single batch.) This improves the quality of the correction by removing one aspect of the technical differences between batches.
```{r}
# Subset all of the samples to the common Ensembl gene annotation
allNames <- unlist(lapply(all.sce, function(x){rownames(x)}))
allNamesNb <- table(allNames)
universe <- names(allNamesNb)[allNamesNb==length(all.sce)]

# Subsetting the SingleCellExperiment object.
uni.sce <- lapply(all.sce, function(x){x[universe,]})
# Also subsetting the variance modelling results, for convenience.
uni.dec <- lapply(all.dec, function(x){x[universe,]})

# Rescale each sample to adjust for differences in sequencing depth between samples
rescaled <- multiBatchNorm(uni.sce, batch = "Sample")

# Compute average variance components across samples
combined.dec <- combineVar(uni.dec)

# Identify highly variables genes
# Here union of hvgs selected by each sample and top 5000 hgvs were selected
all.hvgs.union <- Reduce(union, all.hvgs)
chosen.hvgs <- rownames(combined.dec) %in% union(getTopHVGs(combined.dec, n = 5000), all.hvgs.union)
table(chosen.hvgs)
combined.dec$chosenHvg <- chosen.hvgs
```

When integrating datasets of variable composition, it is generally safer to err on the side of including more genes than are used in a single dataset analysis, to ensure that markers are retained for any dataset-specific subpopulations that might be present. For a top X selection, this means using a larger X (say, ~5000), or in this case, we simply take all genes above the trend.
Alternatively, a more forceful approach to feature selection can be used based on marker genes from within-batch comparisons.

## Diagnosing batch effects
```{r}
# Reminder: the metadata must be the same for each sample for cbind()ing.
# concat matrices:
uncorrected <- do.call(cbind, rescaled)

# copy choice to uncorrected SCE:
all(rownames(combined.dec) == rownames(uncorrected))

rowData(uncorrected) <- cbind(rowData(uncorrected), combined.dec)

# saveRDS(uncorrected, paste0(this.path, "/DataIntegration_uncorrected.RDS"))

# Perform PCA TSNE and UMAP
# Using RandomParam() as it is more efficient for file-backed matrices.
set.seed(1234)
uncorrected <- runPCA(uncorrected,
                      subset_row=chosen.hvgs,
                      BSPARAM=BiocSingular::RandomParam())
uncorrected <- runTSNE(uncorrected, dimred = "PCA")
# uncorrected <- runUMAP(uncorrected, dimred = "PCA")

# saveRDS(uncorrected, paste0(this.path, "/DataIntegration_uncorrected_PCA_TSNE.RDS"))
```

### Clustering
We use graph-based clustering on the components to obtain a summary of the population structure. 

As the samples should be replicates, each cluster should ideally consist of cells from each batch. However, we instead see clusters that are comprised of cells from a single batch. This indicates that cells of the same type are artificially separated due to technical differences between batches.

```{r}
# uncorrected <- readRDS(paste0(this.path, "/DataIntegration_uncorrected_PCA_TSNE.RDS"))

mat.uncorrected <- reducedDim(uncorrected, "PCA")

two.out.uncorrected <- clusterRows(mat.uncorrected, TwoStepParam(
  first = KmeansParam(centers=100),
  second = NNGraphParam(k=2, cluster.fun = "walktrap")),
  full = TRUE)
uncorrected$cluster_twostep <- two.out.uncorrected$clusters

# Get number of cells for each {cluster, batch} pair
clusterTab <- data.frame("clusters" = two.out.uncorrected$clusters, 
                         "batch" = uncorrected$Sample,
                         "group" = uncorrected$cluster)
ClusterInfo <- clusterTab %>% 
  group_by(clusters, batch) %>%
  summarise(cells = n())

ClusterInfo.1 <- clusterTab %>% 
  group_by(clusters, group) %>%
  summarise(cells = n())

p1 <- ggplot(data=ClusterInfo, aes(x=clusters,y=cells, fill=batch)) +
    geom_col() +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  ggtitle("uncorrected, cell numbers") +
  theme(legend.text = element_text(size = 7))
p2 <- ggplot(data=clusterTab, aes(x=clusters, fill = batch)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  ggtitle("uncorrected, proportions") +
  theme(legend.text = element_text(size = 7))
plot_grid(p1, p2, ncol=1)
ggsave(paste0(this.path, "/DataIntegration_uncorrected_batcheffect_barplot.pdf"), width = 9, height = 9)

p1 <- ggplot(data=ClusterInfo.1, aes(x=clusters,y=cells, fill=group)) +
    geom_col() +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  ggtitle("uncorrected, cell numbers") +
  theme(legend.text = element_text(size = 7))
p2 <- ggplot(data=clusterTab, aes(x=clusters, fill = group)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  ggtitle("uncorrected, proportions") +
  theme(legend.text = element_text(size = 7))
plot_grid(p1, p2, ncol=1)
ggsave(paste0(this.path, "/DataIntegration_uncorrected_batcheffect_barplot_group.pdf"), width = 9, height = 9)


# Visualize the uncorrected coordinates using a t-SNE plot
plotTSNE(uncorrected,
          colour_by = "Sample") + 
 facet_wrap(~uncorrected$cluster)
ggsave(paste0(this.path, "/DataIntegration_uncorrected_batcheffect_TSNE_groups.pdf"), width = 14, height = 6)

plotTSNE(uncorrected,
          colour_by = "cluster_twostep") + 
 facet_wrap(~uncorrected$cluster)
ggsave(paste0(this.path, "/DataIntegration_uncorrected_batcheffect_TSNE_clusters.pdf"), width = 14, height = 6)
```

## Mutual Nearest Neighbour correction
```{r}
set.seed(1234)
mnn.out <- fastMNN(rescaled,
                   auto.merge = TRUE,
                   d = 50,
                   k = 20,
                   subset.row = chosen.hvgs,
                   BPPARAM = bp.params)
# saveRDS(mnn.out, file = paste0(this.path, "/DataIntegration_MNN_out.RDS"))
```

```{r}
# mnn.out <- readRDS(paste0(this.path, "/DataIntegration_MNN_out.RDS"))
mnn.out.corre.dim <- dim(reducedDim(mnn.out, "corrected"))
mnn.out.recon.dim <- dim(assay(mnn.out, "reconstructed"))

# add feature selection outcome to mmn.out
# used in other analyses later.
columnsToGet <- setdiff(colnames(combined.dec), "per.block")
combined.dec.df <- combined.dec[,columnsToGet] %>%
  data.frame() %>%
  tibble::rownames_to_column("ID") %>%
  filter(ID %in% rownames(rowData(mnn.out)))
rotationMat <- rowData(mnn.out)$rotation
rowData(mnn.out)$ID <- rownames(rowData(mnn.out))
rowData(mnn.out)$rotation <- NULL

rowData(mnn.out) <- rowData(mnn.out) %>%
  data.frame() %>%
  left_join(combined.dec.df, by="ID") %>%
  DataFrame()

# add rotation back
rowData(mnn.out)$rotation <- rotationMat

# also have gene symbol:
# copied from 'uncorrected'
rowData(mnn.out)$Symbol <- rowData(uncorrected)[rowData(mnn.out)$ID,]$Symbol

# add cell information onto mnn.out
all(rownames(colData(mnn.out)) == rownames(colData(uncorrected)))
colData(mnn.out) <- cbind(colData(mnn.out), colData(uncorrected))

# tidy
rm(columnsToGet, rotationMat)

# saveRDS(mnn.out, file = paste0(this.path, "/DataIntegration_MNN_out_featuresinfo.RDS"))
```

### Clustering
```{r}
mat.mnn <- reducedDim(mnn.out, "corrected")

two.out.mnn <- clusterRows(mat.mnn, TwoStepParam(
  first = KmeansParam(centers=100),
  second = NNGraphParam(k=2, cluster.fun = "walktrap")),
  full = TRUE)
mnn.out$cluster_twostep <- two.out.mnn$clusters


# snn.gr <- buildSNNGraph(mnn.out, use.dimred="corrected", k=20)
# clusters.mnn <- igraph::cluster_walktrap(snn.gr)$membership
# colLabels(mnn.out) <- factor(clusters.mnn)
clusterTab.mnn <- data.frame(clusters=mnn.out$cluster_twostep, batch=mnn.out$Sample, group=mnn.out$cluster)

# Get number of cells for each {cluster, batch} pair
ClusterInfo.mnn <- clusterTab.mnn %>% 
  group_by(clusters, batch) %>%
  summarise(cells = n())

ClusterInfo.mnn.1 <- clusterTab.mnn %>% 
  group_by(clusters, group) %>%
  summarise(cells = n())


mp1 <- ggplot(data=ClusterInfo.mnn, aes(x=clusters,y=cells, fill=batch)) +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
    geom_col() +
    theme(legend.text = element_text(size = 7))
mp2 <- ggplot(data=clusterTab.mnn, aes(x=clusters, fill=batch)) +
  geom_bar(position = "fill") +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.text = element_text(size = 7))

plot_grid(mp1, mp2, ncol=1)
ggsave(paste0(this.path, "/DataIntegration_correctedByMNN_batcheffect_barplot_sample.pdf"), width = 9, height = 9)


mp1.s <- ggplot(data=ClusterInfo.mnn.1, aes(x=clusters,y=cells, fill=group)) +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
    geom_col()
mp2.s <- ggplot(data=clusterTab.mnn, aes(x=clusters, fill=group)) +
  geom_bar(position = "fill") +
    theme(axis.text.x=element_text(angle = 90, hjust = 0)) +
  scale_y_continuous(labels = scales::percent)

plot_grid(mp1.s, mp2.s, ncol=1)
ggsave(paste0(this.path, "/DataIntegration_correctedByMNN_batcheffect_barplot_group.pdf"), width = 9, height = 9)

# dimensinality reduction
set.seed(1234)
mnn.out <- runTSNE(mnn.out, dimred="corrected", BPPARAM = bp.params)
mnn.out <- runUMAP(mnn.out, dimred="corrected", BPPARAM = bp.params)
# saveRDS(mnn.out, file = paste0(this.path, "/DataIntegration_MNN_out_clustering_dimensionreduction.RDS"))

# Visualize the uncorrected coordinates using a t-SNE plot
plotTSNE(mnn.out,
          colour_by = "Sample") + 
 facet_wrap(~mnn.out$cluster)
ggsave(paste0(this.path, "/DataIntegration_correctedByMNN_batcheffect_TSNE_groups.pdf"), width = 14, height = 6)

plotTSNE(mnn.out,
          colour_by = "cluster_twostep") + 
 facet_wrap(~mnn.out$cluster)
ggsave(paste0(this.path, "/DataIntegration_correctedByMNN_batcheffect_TSNE_clusters.pdf"), width = 14, height = 6)

plotTSNE(mnn.out,
          colour_by = "cluster_twostep") + 
 facet_wrap(~mnn.out$Sample)
ggsave(paste0(this.path, "/DataIntegration_correctedByMNN_batcheffect_TSNE_samples.pdf"), width = 24, height = 22)
```


## Report system and packages
```{r}
report(sessionInfo())
```