---
title: "03_marker_gene_identification"
author: ""
date: "2022/10/20"
output: 
  html_document: 
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In order to aid the interpretation of the clustering results, it is helpful to identify genes that are contributing to the separation of cells into those clusters.
The main approach to achieve this, is to identify genes that are differently expressed between clusters. 

## Load library
```{r}
library(scuttle)
library(scater)
library(scran)
library(pheatmap)
library(tidyverse) # always load tidyverse after other packages
```

## Data prepare
We will load two SingleCellExperiment objects generated in previous sections:

The SCE with logcounts normalised within each batch generated in the Batch Correction section.
The SCE with batch-corrected data integration using the batchelor::fastMNN() algorithm, including clusters covered in the Clustering section.

```{r}
# Parameters
this.path <- "../Results/03_marker_gene_identification"
bp.params <- MulticoreParam(workers = 20)
set.seed(1234)
dir.create(this.path)

# normalised within batches without batch-integration correction
uncorrected <- readRDS("../Results/02_normalization_and_dimensionality_reduction_clustering/DataIntegration_uncorrected_PCA_TSNE.RDS")
rownames(uncorrected) <- uniquifyFeatureNames(rownames(uncorrected), rowData(uncorrected)$Symbol)

# data corrected using batch integration with mutual nearest neighbours
corrected <- readRDS("../Results/02_normalization_and_dimensionality_reduction_clustering/DataIntegration_MNN_out_clustering_dimensionreduction.RDS")
rownames(corrected) <- uniquifyFeatureNames(rownames(corrected), rowData(corrected)$Symbol)

# copy cluster assignments to the uncorrected object
# first make sure that the cell names are in the same order
all(colnames(uncorrected) == colnames(corrected))
colData(uncorrected)$cluster_twostep <- factor(colData(corrected)$cluster_twostep)
```

## Identifying cluster marker genes

Although we have defined our clusters based on the batch-corrected expression values, these should not be used for for gene-based analyses like marker gene detection. Instead, we should use the uncorrected (normalised) expression values for differential expression between clusters.

Here, we use the Wilcoxon rank-sum test to perform pairwise comparisons between groups of observations.
```{r}
# Wilcoxon rank-sum test
markers_wilcox_up <- findMarkers(
  uncorrected, 
  groups = uncorrected$cluster_twostep, # clusters to compare
  block = uncorrected$cluster,    # covariates in statistical model
  test.type = "wilcox", 
  direction = "up"
)

markers_wilcox_down <- findMarkers(
  uncorrected, 
  groups = uncorrected$cluster_twostep, # clusters to compare
  block = uncorrected$cluster,    # covariates in statistical model
  test.type = "wilcox",
  direction = "down"
)
# save(markers_wilcox_up, markers_wilcox_down, file = paste0(this.path, "/markers_gene_identification.RData"))
# load(paste0(this.path, "/markers_gene_identification.RData"))

for(i in 1:length(markers_wilcox_up)){
  cluster.path <- paste0(this.path, "/Cluster_", i)
  dir.create(cluster.path)
  markers.up.df <- markers_wilcox_up[[i]]
  plotExpression(uncorrected, 
               features = rownames(markers.up.df)[1:6],
               x = "cluster_twostep",
               colour_by="cluster_twostep")
  ggsave(paste0(cluster.path, "/up_expression_top6_violinplot.pdf"), width = 16, height = 14)
  openxlsx::write.xlsx(markers.up.df, file = paste0(cluster.path, "/up_expression_byWilcoxon_results.xlsx"))
  
  markers.down.df <- markers_wilcox_down[[i]]
  plotExpression(uncorrected, 
               features = rownames(markers.down.df)[1:6],
               x = "cluster_twostep",
               colour_by="cluster_twostep")
  ggsave(paste0(cluster.path, "/down_expression_top6_violinplot.pdf"), width = 16, height = 14)
  openxlsx::write.xlsx(markers.down.df, file = paste0(cluster.path, "/down_expression_byWilcoxon_results.xlsx"))
}


for(i in 1:length(markers_wilcox_up)){
  cluster.path <- paste0(this.path, "/Cluster_", i)
  markers.up.df <- markers_wilcox_up[[i]]
  pdf(paste0(cluster.path, "/up_expression_Top10byAUC_heatmap.pdf"), width = 10, height = 12)
  plotHeatmap(uncorrected, 
            features = rownames(markers.up.df)[markers.up.df$Top <= 10],
            order_columns_by = c("cluster_twostep", "cluster"))
 dev.off()
  
  markers.down.df <- markers_wilcox_down[[i]]
  pdf(paste0(cluster.path, "/down_expression_Top10byAUC_heatmap.pdf"), width = 10, height = 12)
  plotHeatmap(uncorrected, 
            features = rownames(markers.up.df)[markers.up.df$Top <= 10],
            order_columns_by = c("cluster_twostep", "cluster"))
  dev.off()
  
}


```

