---
title: "Differential_Expression_Sliding_Window_Analysis"
author: ""
date: "2021/5/7"
output: 
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

Differential Expression - Sliding Window ANalysis (DE-SWAN) 

## Load library
```{r}
library(DEswan)
library(ggplot2)
library(ComplexHeatmap)
library(tibble)
library(dplyr)
library(pracma)
```

## Data

Age is the quantitative trait of interest and Sex is the covariate we want to include in the analysis

```{r}
load("../../input/omics3_combined/SixHosp_aging_174subjects_omics3_meta_data.RData")
meta <- meta_dat
meta$Gender <- meta$sex
meta$Gender[meta$Gender == "1"] <- "male"
meta$Gender[meta$Gender == "2"] <- "female"
df <- ck_dat
colnames(df) <- make.names(colnames(df), unique = T)
dfs <- log10(df + 0.00001)
# dfs <- df
data <- cbind(meta[,c("age", "Gender")], dfs)
colnames(data)[1:2] <- c("Age", "Sex")

# distribution of qt
hist(data[,1],main="qt distribution",xlab="Age (years)")

# covariates
table(data[,2])
mycols <- as.character(see::metro_colors())
```

## DEswan windown 5
```{r}
x <- cor(data[,1],data[,-c(1:2)])

res.DEswan <- DEswan(data.df = data[,2+c(which(colnames(x) %in% colnames(x)[abs(x)>0]))],
                  qt = data[,1],
                  window.center = seq(20,90,1),
                  buckets.size = 5,
                  covariates = data.frame(data["Sex"]))

# pvalues of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$p)

# coefficients of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$coeff)

# reshape
res.DEswan.wide.p=reshape.DEswan(res.DEswan,parameter = 1,factor = "qt")
DT::datatable(res.DEswan.wide.p)
```

### Pvalues adjustment

The Bonferroni correction is one simple way to take this into account; adjusting the false discovery rate using the Benjamini-Hochberg procedure is a more powerful method.

```{r}
res.DEswan.wide.q <- q.DEswan(res.DEswan.wide.p,method="BH")
DT::datatable(res.DEswan.wide.q)
```

### Calculate and plot the number of significant variables

One simple way to visualize DEswan results is to count the number of significant variables for each window.center tested.

```{r}
res.DEswan.wide.p.signif <- nsignif.DEswan(res.DEswan.wide.p)
toPlot <- res.DEswan.wide.p.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("p<",rownames(toPlot),sep=""),lwd=c(1,2,3))

tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (p < 0.05)") + 
  xlab("Age (years)")



res.DEswan.wide.q.signif <- nsignif.DEswan(res.DEswan.wide.q)
toPlot <- res.DEswan.wide.q.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("q<",rownames(toPlot),sep=""),lwd=c(1,2,3))


tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (q < 0.05)") + 
  xlab("Age (years)")
```



### Heatmap of changes during aging

The best way to identify when and how the features are changing in the qt space is to generate a heatmap of the signed effects.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.coeff <- reshape.DEswan(res.DEswan,parameter = 2,factor = "qt")
toHeatmap <- sign(res.DEswan.wide.coeff[,-1])*-log10(res.DEswan.wide.p[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="row",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```

### Covariates analysis

Similar analysis can be generated for each covariates. Here we show how sex affect the plasma proteome during aging.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.p.covar1 <- reshape.DEswan(res.DEswan,parameter = 1,factor = "Sex")
res.DEswan.wide.coeff.covar1 <- reshape.DEswan(res.DEswan,parameter = 2,factor = "Sexmale")

toHeatmap <- sign(res.DEswan.wide.coeff.covar1[,-1])*-log10(res.DEswan.wide.p.covar1[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff.covar1$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")


gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="both",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```


## DEswan windown 10
```{r}
x <- cor(data[,1],data[,-c(1:2)])

res.DEswan <- DEswan(data.df = data[,2+c(which(colnames(x) %in% colnames(x)[abs(x)>0]))],
                  qt = data[,1],
                  window.center = seq(20,90,1),
                  buckets.size = 10,
                  covariates = data.frame(data["Sex"]))

# pvalues of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$p)

# coefficients of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$coeff)

# reshape
res.DEswan.wide.p=reshape.DEswan(res.DEswan,parameter = 1,factor = "qt")
DT::datatable(res.DEswan.wide.p)
```

### Pvalues adjustment

The Bonferroni correction is one simple way to take this into account; adjusting the false discovery rate using the Benjamini-Hochberg procedure is a more powerful method.

```{r}
res.DEswan.wide.q <- q.DEswan(res.DEswan.wide.p,method="BH")
DT::datatable(res.DEswan.wide.q)
```

### Calculate and plot the number of significant variables

One simple way to visualize DEswan results is to count the number of significant variables for each window.center tested.

```{r}
res.DEswan.wide.p.signif <- nsignif.DEswan(res.DEswan.wide.p)
toPlot <- res.DEswan.wide.p.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("p<",rownames(toPlot),sep=""),lwd=c(1,2,3))

tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (p < 0.05)") + 
  xlab("Age (years)")



res.DEswan.wide.q.signif <- nsignif.DEswan(res.DEswan.wide.q)
toPlot <- res.DEswan.wide.q.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("q<",rownames(toPlot),sep=""),lwd=c(1,2,3))


tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (q < 0.05)") + 
  xlab("Age (years)")
```



### Heatmap of changes during aging

The best way to identify when and how the features are changing in the qt space is to generate a heatmap of the signed effects.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.coeff <- reshape.DEswan(res.DEswan,parameter = 2,factor = "qt")
toHeatmap <- sign(res.DEswan.wide.coeff[,-1])*-log10(res.DEswan.wide.p[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="row",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
pdf("window_10_age_heatmap.pdf", width = 8, height = 14)
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="row",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
dev.off()
```

### Covariates analysis

Similar analysis can be generated for each covariates. Here we show how sex affect the plasma proteome during aging.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.p.covar1 <- reshape.DEswan(res.DEswan,parameter = 1,factor = "Sex")
res.DEswan.wide.coeff.covar1 <- reshape.DEswan(res.DEswan,parameter = 2,factor = "Sexmale")

toHeatmap <- sign(res.DEswan.wide.coeff.covar1[,-1])*-log10(res.DEswan.wide.p.covar1[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff.covar1$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")


gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="both",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
pdf("window_10_sex_heatmap.pdf", width = 8, height = 14)
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="both",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
dev.off()
```



## DEswan windown 15
```{r}
x <- cor(data[,1],data[,-c(1:2)])

res.DEswan <- DEswan(data.df = data[,2+c(which(colnames(x) %in% colnames(x)[abs(x)>0]))],
                  qt = data[,1],
                  window.center = seq(20,90,1),
                  buckets.size = 15,
                  covariates = data.frame(data["Sex"]))

# pvalues of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$p)

# coefficients of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$coeff)

# reshape
res.DEswan.wide.p=reshape.DEswan(res.DEswan,parameter = 1,factor = "qt")
DT::datatable(res.DEswan.wide.p)
```

### Pvalues adjustment

The Bonferroni correction is one simple way to take this into account; adjusting the false discovery rate using the Benjamini-Hochberg procedure is a more powerful method.

```{r}
res.DEswan.wide.q <- q.DEswan(res.DEswan.wide.p,method="BH")
DT::datatable(res.DEswan.wide.q)
```

### Calculate and plot the number of significant variables

One simple way to visualize DEswan results is to count the number of significant variables for each window.center tested.

```{r}
res.DEswan.wide.p.signif <- nsignif.DEswan(res.DEswan.wide.p)
toPlot <- res.DEswan.wide.p.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("p<",rownames(toPlot),sep=""),lwd=c(1,2,3))

tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (p < 0.05)") + 
  xlab("Age (years)")



res.DEswan.wide.q.signif <- nsignif.DEswan(res.DEswan.wide.q)
toPlot <- res.DEswan.wide.q.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("q<",rownames(toPlot),sep=""),lwd=c(1,2,3))


tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (q < 0.05)") + 
  xlab("Age (years)")
```



### Heatmap of changes during aging

The best way to identify when and how the features are changing in the qt space is to generate a heatmap of the signed effects.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.coeff <- reshape.DEswan(res.DEswan,parameter = 2,factor = "qt")
toHeatmap <- sign(res.DEswan.wide.coeff[,-1])*-log10(res.DEswan.wide.p[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="row",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```

### Covariates analysis

Similar analysis can be generated for each covariates. Here we show how sex affect the plasma proteome during aging.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.p.covar1 <- reshape.DEswan(res.DEswan,parameter = 1,factor = "Sex")
res.DEswan.wide.coeff.covar1 <- reshape.DEswan(res.DEswan,parameter = 2,factor = "Sexmale")

toHeatmap <- sign(res.DEswan.wide.coeff.covar1[,-1])*-log10(res.DEswan.wide.p.covar1[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff.covar1$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")


gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="both",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```



## DEswan windown 20
```{r}
x <- cor(data[,1],data[,-c(1:2)])

res.DEswan <- DEswan(data.df = data[,2+c(which(colnames(x) %in% colnames(x)[abs(x)>0]))],
                  qt = data[,1],
                  window.center = seq(20,90,1),
                  buckets.size = 20,
                  covariates = data.frame(data["Sex"]))

# pvalues of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$p)

# coefficients of qT and covariates for each variable and each window.center
DT::datatable(res.DEswan$coeff)

# reshape
res.DEswan.wide.p=reshape.DEswan(res.DEswan,parameter = 1,factor = "qt")
DT::datatable(res.DEswan.wide.p)
```

### Pvalues adjustment

The Bonferroni correction is one simple way to take this into account; adjusting the false discovery rate using the Benjamini-Hochberg procedure is a more powerful method.

```{r}
res.DEswan.wide.q <- q.DEswan(res.DEswan.wide.p,method="BH")
DT::datatable(res.DEswan.wide.q)
```

### Calculate and plot the number of significant variables

One simple way to visualize DEswan results is to count the number of significant variables for each window.center tested.

```{r}
res.DEswan.wide.p.signif <- nsignif.DEswan(res.DEswan.wide.p)
toPlot <- res.DEswan.wide.p.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("p<",rownames(toPlot),sep=""),lwd=c(1,2,3))

tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (p < 0.05)") + 
  xlab("Age (years)")



res.DEswan.wide.q.signif <- nsignif.DEswan(res.DEswan.wide.q)
toPlot <- res.DEswan.wide.q.signif
x <- as.numeric(gsub("X","",colnames(toPlot)))
plot(1, type = "n", xlim=c(min(x,na.rm=T),max(x,na.rm=T)),ylim=c(0,max(toPlot,na.rm=T)),ylab="# significant",xlab="qt")
for(i in 1:nrow(toPlot)){
  lines(x,
        toPlot[i,],type='l',lwd=i)
}
legend("topleft",legend = paste("q<",rownames(toPlot),sep=""),lwd=c(1,2,3))


tmp <- data.frame(toPlot[1,])
tmp <- rownames_to_column(tmp, var = "x")
tmp$x <- gsub("X", "", tmp$x)
colnames(tmp) <- c("x", "y")
tmp$x <- as.numeric(tmp$x)
x_peak <- findpeaks(tmp$y, nups = 1, threshold = 1, sortstr = T)
x_num <- ifelse(nrow(x_peak) < 3, nrow(x_peak), 3)
x_ord <- x_peak[,2][1:x_num]
x_lab <- tmp$x[x_ord]
x_txt <- rep("", nrow(tmp))
x_txt[x_ord] <- x_lab
ggplot(tmp, aes(x, y)) +
  geom_line(size=1) +
  geom_vline(xintercept=x_lab,
                size=1, colour=see::metro_colors(c("blue", "red", "green"))[1:x_num]) +
  geom_text(mapping = aes(x = tmp$x, y = 2, label = x_txt),
                          size=4, angle=90, vjust=-0.4, hjust=0) +
  theme(panel.background = element_rect(fill = "white", colour = "grey50")) +
  ggtitle("Peaks of aging cytokines") +
  ylab("# of significant cytokines (q < 0.05)") + 
  xlab("Age (years)")
```



### Heatmap of changes during aging

The best way to identify when and how the features are changing in the qt space is to generate a heatmap of the signed effects.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.coeff <- reshape.DEswan(res.DEswan,parameter = 2,factor = "qt")
toHeatmap <- sign(res.DEswan.wide.coeff[,-1])*-log10(res.DEswan.wide.p[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")
gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="row",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```

### Covariates analysis

Similar analysis can be generated for each covariates. Here we show how sex affect the plasma proteome during aging.

```{r, fig.height=9, fig.width=5}
res.DEswan.wide.p.covar1 <- reshape.DEswan(res.DEswan,parameter = 1,factor = "Sex")
res.DEswan.wide.coeff.covar1 <- reshape.DEswan(res.DEswan,parameter = 2,factor = "Sexmale")

toHeatmap <- sign(res.DEswan.wide.coeff.covar1[,-1])*-log10(res.DEswan.wide.p.covar1[,-1])
rownames(toHeatmap) <- res.DEswan.wide.coeff.covar1$variable

pairs.breaks <- seq(-3, 3, by=0.01)
mycol <- gplots::colorpanel(n=length(pairs.breaks)-1,low="cyan",mid="black",high="yellow")


gplots::heatmap.2(as.matrix(toHeatmap),
                       cexRow=.5,cexCol=.7,
                       trace="none",
                       dendrogram="both",
                       breaks=pairs.breaks, 
                       col=mycol, 
                       Rowv=T,key=F,
                       Colv=F,
                       lhei=c(0.2,10),
                       lwid=c(.2,3)
)
```

